---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# EFBA

<!-- badges: start -->
[![R-CMD-check](https://github.com/mattByrom-tamu/EFBA/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/mattByrom-tamu/EFBA/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The R package "EFBA" provides functions for implementing empirical band analysis for non stationary time series, with methods for functional and multivariate cases. For reference, see the following.

1: ["Empirical Frequency Band Analysis of Nonstationary Time Series." Bruce, Tang, Hall, Krafty (2020)] (https://www.ncbi.nlm.nih.gov/pmc/articles/PMC8186526/) 

2: ["Adaptive Frequency Band Analysis for Functional Time Series." Bagchi, Bruce (2021)] (https://arxiv.org/abs/2102.01784) 

3: ["Frequency Band Analysis of Nonstationary Multivariate Time Series." Sundararajan, Bruce (2023)] (https://arxiv.org/abs/2301.03664/)

## Installation

You can install the development version of EFBA from [GitHub](https://github.com/mattByrom-tamu/EFBA)

``` r
# install.packages("devtools")
devtools::install_github("mattByrom-tamu/EFBA")
# or
devtools::install_github("mattByrom-tamu/EFBA", build_vignettes=TRUE)
```

## Example

Empirical Band Analysis using the FRESH statistic for linear time-varying dynamics (3 bands). Examples for a white noise (1 band) and sinusoidal time-varying dynamics (3 bands) can be found in the EBA vignette 

```{r example}
library(EFBA)
library(fields)

###Simulate data for all simulated data settings described in paper
set.seed(823819); #if you change the seed, you will get different results
T <- 50000; #total length of time series 
X <- eba.simdata(T=T);

###Run search algorithm using FRESH statistic
##Input parameters
N <- 500; #number of observations per approximately stationary block
K <- 15; #number of tapers to use in multitaper spectral estimator
alpha <- 0.05; #significance level to use for testing partition points using FRESH statistic
std <- FALSE; #should the variance of each stationary block be set to one across all blocks? (TRUE or FALSE)

# Call the search algorithm using the FRESH statistic 
ebaout.bL <- eba.search(X=X$bL,N=N,K=K,std=std,alpha=alpha);

##See final partition of frequency space
ebaout.bL$part.final; #three bands estimated

ebaout.bL$log; #3 partition points tested in total, 2 determined to be significant

image.plot(x=ebaout.bL$mtspec$t,y=ebaout.bL$mtspec$f,z=t(ebaout.bL$mtspec$mtspec), 
           axes = TRUE, col = inferno(256), zlim=c(0,10), 
           xlab='Time',ylab='Hz',xaxs="i"); 
abline(h=ebaout.bL$part.final[c(-1,-length(ebaout.bL$part.final))],col='green');

ebaout.bL$flat 
#p-value for band (0.150,0.344] fails to reject hypothesis of flat spectrum
#p-values for other frequency bands reject flat spectrum hypothesis
```
